### 2022-11-02 charts/index.qmd
Minor

---
title: "Charts"
format: 
  html:
    page-layout: full
editor: visual
---

::: panel-tabset

## 1. Insights in demand for care services

- [1A - Total duration per team by care type](1a.html)
- [1B - Distribution of weekly demand for care services](1b.html)
- [1C - Total demand for care per day by part of days](1c.html)

## 2. Demand vs supply

## 3. Continuity of service

## 4. Scale / size teams

:::