### 2023-10-25 paper-presentation/paper-presentation.qmd
Final version

---
title: "Team size and composition in home health care"
author: "René Bekker, Yoram Clapper, Witek ten Hove and Dennis Moeke"
format:
  revealjs:
    logo: images/logo-h.png
    navigation-mode: vertical
    theme: [default, custom.scss]
bibliography: references.bib
execute:
  output: true
jupyter: python3
---

# Social issue

![@ministerie_van_volksgezondheid_home_2020](images/supply.png)

## Social issue

```{python}
import pandas as pd
import plotly.graph_objects as go

# Read the CSV file
df = pd.read_csv('absenteeism.csv')
df

# Create the line chart
fig = go.Figure()

# Add the trace for the data
fig.add_trace(go.Scatter(x=df['periode'], y=df['verzuim'], mode='lines', line_color='tomato', name='Absenteeism Percentages'))

# Title and labels (translated to English)
fig.update_layout(
    title='Absenteeism Percentages Dutch Healthcare Sector<br><sub>Annual Figures (%)</sub>',
    xaxis_title='Period',
    yaxis_title='Percentage',
    annotations=[
        dict(
            x=1,
            y=-0.2,
            xref='paper',
            yref='paper',
            text='Source: CBS',
            showarrow=False
        )
    ]
)

# Display the chart
fig.show()
```

## Decision problem

![From: @grieco2021operational](decisionproblems.png)

## Business problem

-   What are demand dynamics?
-   What are appropriate team sizes?

# Principles

[1 - Care demand]{style="color: red;"}

2 - Travel time

[3 - Effective capacity]{style="color: red;"}

[4 - Team composition]{style="color: red;"}

5 - Contract type

[6 - Communication and management]{style="color: red;"}

# Data understanding

![](images/database.png)

## Data understanding

| Area type | Teams | Planned care \n(hrs/wk) |    Clients | Clients/km^2^ |
|-----------|------:|------------------------:|-----------:|--------------:|
| Urban     |    26 |                238 (51) | 221 (51.4) |     71 (31.4) |
| Suburban  |    18 |                224 (54) | 238 (61.3) |      11 (5.4) |
| Rural     |    11 |                213 (46) | 176 (53.6) |       3 (1.1) |

: General information per area type for the years 2020, 2021; all numerical values (except team count) represent the mean, with standard deviation in brackets.

## Data understanding

![](images/durationperday.png)

# Data preparation

![](images/meansvar.png)

## Data preparation

![](images/los.png)

# Modeling

# Principle 1: Care Demand

## Principle 1: Care Demand

::: columns
::: {.column width="50%"}
Demand per week [$N$]{style="color: red;"} is generated by:

-   Arrival rates [$a$]{style="color: red;"}: number of new cases per week
-   Length of stays [$s$]{style="color: red;"}: number of weeks per case
-   Amount of care [$g$]{style="color: red;"}: hours per case per week
:::

::: {.column width="50%"}
![](images/demand.png)
:::
:::

## Principle 1: Care Demand

For some grade of service [$\gamma$]{style="color: red;"}, the required weekly capacity [$C$]{style="color: red;"} is $$C = \rho + \gamma z \rho$$

where [$\rho$]{style="color: red;"} $= m_am_sm_g$ is the average demand, and [$z$]{style="color: red;"} is the peakedness given by $$z = \frac{\sigma_g^2}{m_g} + m_g + m_g(1−G_s)(\frac{\sigma_a^2}{m_a}-1)$$
Hence, the utilization of the capacity [$C$]{style="color: red;"} is $$\frac{\mathbb{E}[N]}{C} = \frac{1}{1 + \frac{\gamma\sqrt{z}}{\sqrt{\rho}}}$$

## Principle 1: Care Demand

::: columns
::: {.column width="40%"}
*The buffer capacity required to handle demand variability decreases with scale, but the possible reduction becomes smaller as the scale increases.*

[NB: The available data is limited to planned care and may not accurately reflect the actual demand for home care.]{style="color: red;"}
:::

::: {.column width="60%"}
```{python}
#------------------------------------------------------------------------------
# Principle 1: care demand
#------------------------------------------------------------------------------

# generates image of utility vs average weekly demand
# gamma: grade of service, typically between [0.5, 2.]
# z: variance-to-mean ratio (variance weekly demand / mean weekly demand)
# max_rho: upper limit of x-axis (maximum weekly demand in hours to observe)
# min_rho: lower limit of x-axis (maximum weekly demand in hours to observe)
import matplotlib.pyplot as plt 
import numpy as np 
import scipy.integrate as integrate 
from scipy.stats import norm
import networkx as nx

def plot_care_demand(gamma,idx,z,d,min_rho,max_rho):
    fig, ax = plt.subplots(figsize=(9.6,7.2))
    ax.set_xlabel('Average weekly demand in hours',size=12)
    ax.set_ylabel('Utilization of capacity',size=12)

    rho = np.arange(min_rho, max_rho)
    N = len(idx)
    color = ['blue','orange','green']
    for i in range(N):
        nr = idx[i]
        z_i = z[i]
        d_i = d[i]
        u_i = 1 / (1 + gamma*np.sqrt(z_i/rho))
        c_i = color[i]
        ax.plot(rho,u_i,color=c_i,label='Team '+str(nr),linestyle='solid')

        xlim = ax.get_xlim()
        ylim = ax.get_ylim()
        ax.set_xlim(xlim)
        ax.set_ylim(ylim[0],1)

        pin_u = 1 / (1 + gamma*np.sqrt(z_i/d_i))
        ax.hlines(pin_u,xmin=xlim[0],xmax=d_i,color='grey',linestyle='dashed')
        ax.vlines(d_i,ymin=ylim[0],ymax=pin_u,color='grey',linestyle='dashed')
        ax.plot(d_i, pin_u,color=c_i,marker='s')

    ax.legend()
    plt.show()  # This will display the plot
    plt.close()

plot_care_demand(gamma=1,idx=[5,32,29],z=[6.99,9.01,18.49],d=[132.30,370.87,231.59],min_rho=50,max_rho=1000)
```
:::
:::

# Principle 3: Effective capacity

## Principle 3: Effective capacity

![](images/workingtime.png)

## Principle 3: Effective capacity

Let [$M$]{style="color: red;"} be the total number of scheduled care workers during some period [$T$]{style="color: red;"}, and let [$p$]{style="color: red;"} be the probability that the care worker is present.

The number of care workers present [$\tilde{M}$]{style="color: red;"} then follows a Binomial($M$, $p$) distribution. Consequently, the properties of the effective capacity

[$P$]{style="color: red;"} $= \frac{\tilde{M}}{M}$

follow directly from this observation.

## Principle 3: Effective capacity

For $k\in [0,1]$, the probability that the effective capacity [$P$]{style="color: red;"} is at most [$k$]{style="color: red;"} equals

$$\mathbb{P}(P \le k) = \mathbb{P}(\tilde M \le k M)
 = \frac{B(kM+1,(1-k)M,p)}{B(kM+1,(1-k)M)}$$ with $B(x,y,p) = \int_p^1 t^{x-1} (1-t)^{y-1} {\rm d}t$ the incomplete Beta function and $B(x,y) = B(x,y,0)$.

## Principle 3: Effective capacity

*Small teams are more prone to lower levels of effective capacity than large teams as a result of variability in leave of absence and sick leave, whereas the differences between larger teams become smaller.*

```{python}
#------------------------------------------------------------------------------
# Principle 3: productivity
#------------------------------------------------------------------------------
# calculates cdf of productivity: Pr(P <= x)
# M: total scheduled workers
# p: probability a worker is present on scheduled moment
def productivity_cdf(x,M,p):

    # functions to calculate incomplete beta function
    def poly(t,x,y) : return t**(x-1) * (1-t)**(y-1)
    def betainc(x,y,pr) : return integrate.quad(lambda t : poly(t,x,y), pr, 1)

    cdf = betainc(x*M + 1, (1 - x)*M, p)[0] / betainc(x*M + 1, (1 - x)*M, 0)[0]

    return cdf

# generates image of productivity cdf vs total scheduled workers
# targets(list): all (cdf) values to be evaluated: i.e. all x's to Pr(P <= x)
# p: probability a worker is present on scheduled moment
# max_M: upper limit of x-axis (maximum scheduled workers)
# min_M: lower limit of x-axis (minimum scheduled workers) 
def plot_productivity(targets,p,min_M,max_M,pinned_size):
    fig, ax = plt.subplots(figsize=(8,5))
    M = np.arange(min_M, max_M)

    if type(targets) == float or type(targets) == int : targets = [targets]
    for x in targets:
        label = 'k = '+str(x)
        cdf = np.array([productivity_cdf(x,m,p) for m in M])
        if x == 0.6:
            xcdf = cdf.copy()
        ax.plot(M,cdf,label=label)

    xlim = ax.get_xlim()
    ylim = ax.get_ylim()
    ax.set_xlim(xlim)
    ax.set_ylim(ylim)

    pinned_cdf = xcdf[pinned_size-1]
    ax.hlines(pinned_cdf,xmin=xlim[0],xmax=pinned_size,color='grey',linestyles='dashed')
    ax.vlines(pinned_size,ymin=ylim[0],ymax=pinned_cdf,color='grey',linestyles='dashed')
    ax.plot(pinned_size, pinned_cdf,color='orange',marker='o')

    pinned_size_db = 19
    pinned_cdf = xcdf[pinned_size_db-1]
    ax.hlines(pinned_cdf,xmin=xlim[0],xmax=pinned_size_db,color='grey',linestyles='dashed')
    ax.vlines(pinned_size_db,ymin=ylim[0],ymax=pinned_cdf,color='grey',linestyles='dashed')
    ax.plot(pinned_size_db, pinned_cdf,color='orange',marker='o')

    #ax.set_title('Productivity as a function of team size',size=16)
    ax.set_xlabel('Team size',size=12)
    ax.set_ylabel('$Pr(P \leq k)$',size=12)
    ax.legend()

    plt.show()  # This will display the plot
    plt.close()

plot_productivity([0.5,0.6,0.7], p=0.79, min_M=1, max_M=40, pinned_size = 7)
```

# Principle 4: Team composition

## Principle 4: Team composition

| Level | Qualification | Type of tasks | \% of planned care |
|------:|---------------|---------------|-------------------:|
|     1 | PV niveau 2+  | Personal Care |                67% |
|     2 | PV niveau 3   | Personal Care |                 9% |
|     3 | VP niveau 3   | Nursing       |                24% |

## Principle 4: Team composition

![](images/3a.png)

## Principle 4: Team composition

::: columns
::: {.column width="75%"}

```{python}
#------------------------------------------------------------------------------
# Principle 4: team composition
#------------------------------------------------------------------------------

# computes the expected excess for available capacity under assumption of normal distribution
# cap (float): available capacity
# mean: mean weekly demand (normal distribution)
# std: standard deviation of weekly demand (normal distribution)
def expected_excess(cap,mean,std):
    c = (cap-mean)/std
    s = std
    return s*c*norm.cdf(c) + (s/np.sqrt(2))*np.exp(-0.5*c**2)

# returns list of optimal capacity per qualification level obtained recursively
# maxQ (int): highest qualification level (ranked by integer values)
# mu (list): mean weekly demand per qualification level
# std (list): standard deviation of weekly demand per qualification level
# z (float): service level (based on normal distribution)
# NOTE: all lists need to be ordered by rank of qualification levels
def get_capacity(maxQ,mu,sigma,z):
    k = maxQ + 0
    capacity = np.zeros(maxQ)
    excess = np.zeros(maxQ)
    while k > 0:
        m_k = mu[k-1]
        s_k = sigma[k-1]
        e_k = excess[k-1]

        v = m_k + z*s_k - e_k
        if v > 0:
            capacity[k-1] = v

        k -= 1
        val = np.sum(capacity[k:] - excess[k:])
        mean = np.sum(mu[k:])
        std = np.sqrt(np.sum(sigma[k:]**2))
        excess[k-1] = expected_excess(val,mean,std)
    return capacity

# generates image of optimal capacity distribution over qualification level
# maxQ (int): highest qualification level (ranked by integer values)
# mu (list): mean weekly demand per qualification level
# sigma (list): standard deviation of weekly demand per qualification level
# z (float): service level (based on normal distribution)
# max_rate: upper limit x-axis (scaling by arrival rate)
# min_rate: lower limit x-axis (scaling by arrival rate)
def plot_optimal_cap_on_qlevel(maxQ,mu,sigma,z,min_rate,max_rate):
    fig, ax = plt.subplots(figsize=(14,6.8))
    mu = np.array(mu)
    sigma = np.array(sigma)

    rate = np.arange(min_rate,max_rate+0.1,0.1)
    norm_cap_matrix = np.zeros((len(rate),maxQ))
    for i, m in enumerate(rate):
        mu_scaled = m * mu
        sigma_scaled = np.sqrt(m) * sigma
        capacity = get_capacity(maxQ,mu_scaled,sigma_scaled,z)
        cap_sum = np.sum(capacity)
        norm_cap_matrix[i] = capacity/cap_sum

    mean_norm = mu / np.sum(mu)
    color = {0:'blue',1:'red',2:'green',3:'orange'}
    label = {0:'PV niveau 2+',1:'PV niveau 3',2:'VP niveau 3'}
    for k in np.arange(maxQ):
        capx = norm_cap_matrix.T[k]
        meanx = np.array([mean_norm[k] for _ in range(len(rate))])
        ax.plot(rate,capx,c=color[k],label=label[k])
        ax.plot(rate,meanx,c=color[k],linestyle='dashed')
    ax.legend()
    #ax.set_title('Optimal distribution of capacity over qualification levels',size=16)
    ax.set_xlabel('$m_a$ (scaling factor)',size=12)
    ax.set_ylabel('Capacity ratio',size=12)

    limx = ax.get_ylim()
    ax.vlines(1, ymin=limx[0], ymax=limx[1], color='grey', linestyles='dashed')
    fig.text(.5, .01, 'Ratio in capacity $C_k/ ∑_{j∈Q} C_j$ as mean number of new clients $m_a$ increases; dashed horizontal lines represent ratios in demand $μ_k/ ∑_{j∈Q} μ_j$. Dashed vertical line corresponds to current demand.', ha='center')
    plt.show()  # This will display the plot
    plt.close()

plot_optimal_cap_on_qlevel(maxQ=3, mu=[49.16, 20, 63.13], sigma=[12.85, 9.29, 17.92], z=1, min_rate=0.1, max_rate=5)
```
:::

::: {.column width="25%"}
![](images/3acrop.png)
:::
:::

## Principle 4: Team composition

*Small teams must deploy above-average amounts of high-level care workers to sufficiently cope with the variability in demand. As the scale increases, the amount of capacity required will move closer to the average workload for each qualification level.*

# Principle 6: Communication and management

## Principle 6: Communication and management

::: columns
::: {.column width="60%"}
Let [$M$]{style="color: red;"} be the number of members of a team. Then there are $$\frac{M(M-1)}{2}$$ potential interactions between team members. In terms of complexity the number of interactions between a team of size [$M$]{style="color: red;"} equals $O(M^2)$
:::

::: {.column width="40%"}
```{python}
def plot_complete_graph(M):
    plt.clf()
    # Create a complete graph with M nodes
    G = nx.complete_graph(M)

    # Draw the graph
    pos = nx.spring_layout(G)  # positions for all nodes
    nx.draw_networkx_nodes(G, pos, node_size=700, node_color="tomato")
    nx.draw_networkx_edges(G, pos, width=1.0, alpha=0.5)
    nx.draw_networkx_labels(G, pos, font_size=20, font_color="white")

    plt.title("Complete Graph with {} nodes".format(M))
    plt.show()
    plt.close()

# Test the function
plot_complete_graph(5)
```
:::
:::

## Principle 6: Communication and management

```{python}
import matplotlib.pyplot as plt
import networkx as nx

def plot_graph(N, k, M):
    plt.clf()
    G = nx.Graph()

    if N == 0:
        # Create a complete graph with M nodes
        G = nx.complete_graph(M)

    elif N == 1:
        central_node = 0
        G.add_node(central_node)

        nodes_per_graph = M // k
        for i in range(k):
            # Create a complete graph for each k
            start = i * nodes_per_graph + 1
            end = start + nodes_per_graph
            subgraph = nx.complete_graph(range(start, end))

            # Add the subgraph to the main graph
            G = nx.union(G, subgraph)

            # Connect the central node to one node in each subgraph
            G.add_edge(central_node, start)

    # Compute the number of nodes and edges
    total_nodes = G.number_of_nodes()
    total_edges = G.number_of_edges()

    # Draw the graph
    pos = nx.spring_layout(G)
    nx.draw_networkx_nodes(G, pos, node_color="tomato")
    nx.draw_networkx_edges(G, pos)
    nx.draw_networkx_labels(G, pos, font_color="white")

    # Set the title
    plt.title(f"Graph with {total_nodes} nodes and {total_edges} edges")

    plt.show()
    plt.close()
```

::: columns
::: {.column width="50%"}
```{python}
plot_graph(0, 3, 12);  # Complete graph with 12 nodes
```
:::

::: {.column width="50%"}
```{python}
plot_graph(1, 3, 12);  # 3 complete graphs with 2 nodes each connected to a central node
```
:::
:::

## Principle 6: Communication and management

*The complexity of the number of interactions for a team of size* [$M$]{style="color: red;"} is $O(M^2)$. By splitting the team into [$k$]{style="color: red;"} sub-teams, managed by one mediator, the number of interactions can be reduced by a factor [$k$]{style="color: red;"} as [$M$]{style="color: red;"} grows large.

```{python}
#------------------------------------------------------------------------------
# Principle 6: communication and management
#------------------------------------------------------------------------------

# generates image of number of interactions per team size under splitted teams
# team_splits: maximum number of times a team can be (equally) splitted
# max_team_size: upper limit x-axis (largest team size to observe)
def plot_interactions(team_splits,max_team_size, pinned_size):
    fig, ax = plt.subplots()

    sizes = np.arange(max_team_size+1)
    for k in range(1,team_splits+1):
        inter = 0.5 * sizes * (1 + (1/k) * sizes)
        if k == 1:
            ax.plot(sizes,inter,label='Complete team')

            xlim = ax.get_xlim()
            ylim = ax.get_ylim()
            ax.set_xlim(xlim)
            ax.set_ylim(ylim)

            pinned_target = 0.5 * pinned_size * (1 + (1/k) * pinned_size)
            ax.hlines(pinned_target, xmin=xlim[0], xmax=pinned_size, color='grey', linestyles='dashed')
            ax.vlines(pinned_size, ymin=ylim[0], ymax=pinned_target, color='grey', linestyles='dashed')
            ax.plot(pinned_size, pinned_target, 'bo')

            pinned_size_db = 19
            pinned_target_db = 0.5 * pinned_size_db * (1 + (1/k) * pinned_size_db)
            ax.hlines(pinned_target_db, xmin=xlim[0], xmax=pinned_size_db, color='grey', linestyles='dashed')
            ax.vlines(pinned_size_db, ymin=ylim[0], ymax=pinned_target_db, color='grey', linestyles='dashed')
            ax.plot(pinned_size_db, pinned_target_db, 'bo')

        else:
            ax.plot(sizes,inter,label='Team split in '+str(k))

            if k == 2:
                xlim = ax.get_xlim()
                ylim = ax.get_ylim()
                ax.set_xlim(xlim)
                ax.set_ylim(ylim)

                pinned_target = 0.5 * pinned_size * (1 + (1/k) * pinned_size)
                ax.hlines(pinned_target, xmin=xlim[0], xmax=pinned_size, color='grey', linestyles='dashed')
                ax.vlines(pinned_size, ymin=ylim[0], ymax=pinned_target, color='grey', linestyles='dashed')
                ax.plot(pinned_size, pinned_target, color = 'orange', marker = 'o')

                pinned_size_db = 19
                pinned_target_db = 0.5 * pinned_size_db * (1 + (1/k) * pinned_size_db)
                ax.hlines(pinned_target_db, xmin=xlim[0], xmax=pinned_size_db, color='grey', linestyles='dashed')
                ax.vlines(pinned_size_db, ymin=ylim[0], ymax=pinned_target_db, color='grey', linestyles='dashed')
                ax.plot(pinned_size_db, pinned_target_db, color = 'orange', marker = 'o')

    ax.legend()
    plt.xticks(range(0,max_team_size+1,5))
    ax.set_xlabel('Team size',size=12)
    ax.set_ylabel('Number of interactions',size=12)

    plt.show()  # This will display the plot
    plt.close()

plot_interactions(team_splits=4, max_team_size=20, pinned_size=7);
```

# Practice based scenarios

::: columns
::: {.column width="35%"}
![](images/map.png){width="300"}
:::

::: {.column width="65%"}
![](images/scenarios.png){width="900"}
:::
:::

# Questions

# References {.unnumbered}

::: {#refs}
:::